---
title: "R Notebook"
output: html_notebook
---

#Packages
```{r}
library(tidyverse)
library(ggplot2)
```

#data
```{r}
overseas_purpose_of_visit <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/overseas_purpose_of_visit_clean.csv") 

overseas_residents_uk_visits <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/overseas_residents_uk_visits_clean.csv") 

spend_per_day_by_purpose <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/overseas_purpose_spend_per_day_vs_visit_clean.csv") 

spend_by_region <- read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/overseas_spending_region_clean.csv") 

spend_by_mode_of_transport_and_purpose <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/spending_by_mode_of_travel_and_purpose.csv") 

visits_and_spending_by_purpose_and_length_of_stay_per_region <- read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_and_spending_by_purpose_and_length_of_stay_per_region.csv") 

visits_and_spending_by_purpose_and_length_of_stay <- read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_and_spending_by_purpose_and_length_of_stay.csv")

visits_to_and_spending_in_uk_regions <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_to_and_spending_in_uk_regions.csv")

visits_to_uk_regions_by_qtr_2018_22 <-
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_to_uk_regions_by_qtr_2018_22")

mode_of_travel_and_purpose <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/mode_of_travel_and_purpose.csv")

visits_to_and_spending_in_uk_regions_by_purpose <-  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_to_and_spending_in_uk_regions_by_purpose.csv")

visits_by_country_2018_22 <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/visits_by_country_2018_22")

spend_by_country_2018_22 <- 
  read_csv("~/Desktop/projects/visit_scotland_project/data/clean_data/spend_by_country_2018_22.csv")
```
```{r}
visit_scotland_colours <- c("#533D7E", "#AF7AD7", "#2C1572", "#6F73D2")
col_scheme <- c("#533D7E", "#AF7AD7", "#2C1572", "#6F73D2")

```

```{r}
# Show the total number of visits for each year by region
overseas_residents_uk_visits_annual <- overseas_residents_uk_visits %>% 
  group_by(year, region) %>% 
  summarise("total_number_of_visits_thousands" = sum(number_of_visits_thousands))
```

```{r}
overseas_residents_uk_visits_annual %>% 
  filter(region != "all_visit") %>% 
  ggplot(aes(x = year, 
             y = total_number_of_visits_thousands, 
             fill = region)) +
  geom_col() +
  labs(title = "Annual UK Visits by Region",
       x = "Year",
       y = "Number of Visits (Thousands)")
```

Here we can see the spread of regions people tavel to the UK from. The UK recieved 
approximately 40 million visitors from overseas in 2018 and showed a slight increase 
in 2019. However we can see a sharp decrease in 2020/ 20201 due to the covid pandemic 
there has been a significant increase in overseas visitor numbers reaching just over 
30 million however the UK is yet to return to pre pandemic levels. The majority of 
overseas visits are from Europe. 


## Purpose of visits
```{r}
# Show the total number of visits for each year by purpose 
overseas_purpose_of_visit_annual <- overseas_purpose_of_visit %>% 
  group_by(year, purpose_of_visit) %>% 
  summarise("total_number_of_visits_per_purpose_thousands" = sum(number_of_visits_thousands),
            "annual_total_number_of_visits_thousands" = sum(all_visits))
```


```{r}
overseas_purpose_of_visit_annual %>% 
  ggplot(aes(x = year, 
             y = total_number_of_visits_per_purpose_thousands, 
             fill = purpose_of_visit )) +
  geom_col() + 
  labs(title = "Annual Purpose of Oversea Visits",
       x = "Year",
       y = "Number of Visits (Thousands)")
```
```{r}
overseas_purpose_of_visit %>% 
  group_by(, year) %>% 
  summarise(sum(total_number_of_visits_per_purpose_thousands))
```

#Spend per day by Purpose

```{r}
spend_per_day_by_purpose %>% 
  ggplot(aes(x = year,
             y = avg_spending_per_visit,
             fill = purpose)) +
  geom_col() +
  labs(title = "Average Spending per Visit by Purpose 2018-2022",
       x = "Year",
       y = "Average Spending per visit (£)") +
  coord_flip()
```

```{r}
spend_per_day_by_purpose %>% 
  ggplot(aes(x = year,
             y = avg_spending_per_day,
             fill = purpose)) +
  geom_col() +
  labs(title = "Average Spending per day by Purpose 2018-2022",
       x = "Year",
       y = "Average Spending per day (£)") 
```

```{r}
spend_per_day_by_purpose %>% 
  ggplot(aes(x = region,
             y = avg_spending_per_day,
             fill = purpose)) +
  geom_col() +
  labs(title = "Average Spending per Visit by Region and Purpose",
       x = "Region",
       y = "Average Spending per visit (£)") +
  coord_flip()
```

```{r}
spend_per_day_by_purpose %>% 
  ggplot(aes(x = region,
             y = avg_spending_per_day,
             fill = purpose)) +
  geom_col() +
  labs(title = "Average Spending per Day by Region and Purpose",
       x = "Region",
       y = "Average Spending per Day (£)") +
  coord_flip()
```

# Spend per day by Region 

```{r}
spend_by_region %>% 
  ggplot(aes(x= year,
             y = spending_million,
             fill = region)) +
  geom_col() +
  labs(title = "Spending by Region",
       x = "Year",
       y = "Spending (£ Million)")
```

```{r}
spend_by_region %>% 
  ggplot(aes(x= region,
             y = spending_million,
             fill = purpose)) +
  geom_col() +
  labs(title = "Spending by Region and Purpose ",
       x = "Region",
       y = "Spending (£ Million)") +
  coord_flip()
```
 # Spend by mode of transport 
```{r}
spend_by_mode_of_transport_and_purpose %>% 
  ggplot(aes(x= year,
             y = total_spending_million,
             fill = mode_of_travel)) +
  geom_col() +
  labs(labs = "Total Spending by Mode of Transport", 
       x = "Year",
       y = "Spending Million (£)")
  
```
 
# Visits by mode of trannsport 

```{r}
selected_colors <- c(col_scheme[3], col_scheme[4]) 

mode_of_travel_and_purpose %>% 
  filter(mode_of_travel != "Total") %>% 
    ggplot(aes(x= year,
             y = number_of_visits,
             fill = mode_of_travel)) +
  geom_col() +
  labs(title = "Total Number of Visits by Mode of Transport", 
       x = "Year",
       y = "Number of Visits") +
    theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm")) +
  scale_fill_manual(values = selected_colors)
  
```
```{r}
travel_graph <- mode_of_travel_and_purpose %>%
  select(-c(purpose, average_annual_growth_2018_22_percent)) %>%
  filter(mode_of_travel != "Total",
         year != "2020") %>%
  group_by(mode_of_travel, year) %>%
  summarise(total_visits = sum(number_of_visits), .groups = "drop")
```

```{r}
selected_colors <- c(col_scheme[3], col_scheme[4]) 

travel_graph %>% 
  ggplot(aes(x = year,
             y = total_visits,
           color = mode_of_travel)) +
  geom_line() +
  labs(title = "Total Number of Visits by Mode of Transport", 
       x = "Year",
       y = "Number of Visits") +
    theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm")) +
  scale_colour_manual(values = selected_colors)
```


```{r}


travel_graph%>%
  ggplot(aes(x = year),
             y = number_of_visits,
             colour = mode_of_travel)) +
  geom_line() +
  labs(title = "Total Number of Visits by Mode of Transport",
       x = "Year",
       y = "Number of Visits") +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.5, "cm")
  ) 
```

## visits_and_spending_by_purpose_and_length_of_stay
```{r}
visits_and_spending_by_purpose_and_length_of_stay %>% 
  glimpse()
```
```{r}
visits_and_spending_by_purpose_and_length_of_stay %>% 
  distinct(length_of_stay)
```

```{r}
visits_and_spending_by_purpose_and_length_of_stay <- 
  visits_and_spending_by_purpose_and_length_of_stay %>% 
  mutate(length_of_stay = recode(length_of_stay, 
                          "30 nights or more" = "29 nights or more",
                          "31 nights or more" = "29 nights or more",
                          "32 nights or more" = "29 nights or more",
                          "33 nights or more" = "29 nights or more",
                          "34 nights or more" = "29 nights or more"))
```


```{r}
visits_and_spending_by_purpose_and_length_of_stay %>% 
  ggplot(aes(x = length_of_stay,
             y = number_of_visits,
             fill= purpose)) +
  geom_col() +
  labs(title = "Number of Visits per Length of Stay",
       x = "Length of Stay",
       y = "Number of Visits") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```
```{r}
visits_and_spending_by_purpose_and_length_of_stay %>% 
  ggplot(aes(x = length_of_stay,
             y = spending_million,
             fill= purpose)) +
  geom_col() +
  labs(title = "Spending vs Length of Stay",
       x = "Length of Stay",
       y = "Spending Million (£)") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```

```{r}
visits_and_spending_by_purpose_and_length_of_stay %>% 
  ggplot(aes(x = year,
             y = spending_million,
             fill = length_of_stay)) +
  geom_col() +
  labs(title = "Spending vs Length of Stay",
       x = "Length of Stay",
       y = "Spending Million (£)") +
  scale_y_continuous(labels = scales::comma) 
```

## visits_and_spending_by_purpose_and_length_of_stay_per_region

```{r}
visits_and_spending_by_purpose_and_length_of_stay_per_region <- visits_and_spending_by_purpose_and_length_of_stay_per_region %>% 
  mutate(length_of_stay = recode(length_of_stay, 
                          "30 nights or more" = "29 nights or more",
                          "31 nights or more" = "29 nights or more",
                          "32 nights or more" = "29 nights or more",
                          "33 nights or more" = "29 nights or more",
                          "34 nights or more" = "29 nights or more",
                          "35 nights or more" = "29 nights or more"))
  
```

```{r}
visits_and_spending_by_purpose_and_length_of_stay_per_region %>% 
  ggplot(aes(x = length_of_stay,
             y = number_of_visits,
             fill= region)) +
  geom_col() +
  labs(title = "Number of Visits per Length of Stay",
       x = "Length of Stay",
       y = "Number of Visits") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```
```{r}
visits_and_spending_by_purpose_and_length_of_stay_per_region %>% 
  ggplot(aes(x = length_of_stay,
             y = spending_million,
             fill= region)) +
  geom_col() +
  labs(title = "Spending per Length of Stay",
       x = "Length of Stay",
       y = "Spending Million (£)") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip()
```
# visits_to_and_spending_in_uk_regions

```{r}
visits_to_and_spending_in_uk_regions %>% 
  ggplot(aes(x = year,
             y = number_of_visits_thousands,
             fill = region_vistited)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Number of Visits by Region",
       x = "Year",
       y = "Number of Visits (Thousands)")
  
```

```{r}
visits_to_and_spending_in_uk_regions %>% 
  ggplot(aes(x = year,
             y = spending_million,
             fill = region_vistited)) +
  geom_col() +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Spending by Region",
       x = "Year",
       y = "Spending (£ Million)")
  
```

```{r}
visits_to_and_spending_in_uk_regions %>% 
  filter(region_vistited == "Scotland") %>% 
  ggplot(aes(x = year,
             y = number_of_visits_thousands)) +
  geom_col(fill = visit_scotland_colours[1]) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Number of Visits to Scotland",
       x = "Year",
       y = "Number of Visits (Thousands)") 
  
```

```{r}
visits_to_and_spending_in_uk_regions %>% 
  filter(region_vistited == "Scotland") %>% 
  ggplot(aes(x = year,
             y = spending_million)) +
  geom_col(fill = visit_scotland_colours[2]) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Spending in Scotland",
       x = "Year",
       y = "Spending (£ Million)") + 
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm")) +
  scale_colour_manual(values = visit_scotland_colours)
  
```

#
```{r}
visits_to_uk_regions_by_qtr_2018_22 %>% 
  ggplot(aes(x= year,
             y= number_of_visits_thousands,
             fill = region_visited)) +
  geom_col() +
  labs(title = "Number of Visits by Region per Quarter",
      x = "Year",
      y = "Number of Visits (Thousands)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
visits_to_uk_regions_by_qtr_2018_22 %>% 
  filter(region_visited == "Scotland") %>% 
  ggplot(aes(x= year,
             y= number_of_visits_thousands,
             fill = region_visited)) +
  geom_col(fill = visit_scotland_colours[1]) +
  labs(title = "Number of Visits to Scotland per Quarter",
      x = "Year",
      y = "Number of Visits (Thousands)") +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm"))
```
```{r}
visits_to_and_spending_in_uk_regions_by_purpose <- visits_to_and_spending_in_uk_regions_by_purpose %>% 
  mutate(purpose = recode(purpose,
                           "of which inclusive tour" = "Holiday"))
```

```{r}
purpose_colours <- c("#AED6F1", "#C39BD3", "#B2BABB","#533D7E")

purpose_visits <- visits_to_and_spending_in_uk_regions_by_purpose %>% 
  filter(region == "Scotland",
         purpose != "All visits") %>% 
  mutate(number_of_visits = number_of_visits_thousands*1000)
purposes <- c("Holiday", "Visiting friends or relatives", "Business","Miscellaneous")
purpose_visits$purpose <- factor(purpose_visits$purpose, levels = purposes, ordered = TRUE)


ggplot(purpose_visits, aes(x = year,
           y = number_of_visits,
           fill = purpose)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
    labs(title = "Number of Visits to Scotland Annually by Purpose",
      x = "Year",
      y = "Number of Visits") +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm")) +
  scale_fill_manual(values = purpose_colours)
  
```


```{r}
purpose_colours <- c("#AED6F1", "#C39BD3", "#B2BABB","#533D7E")

purpose_visits <- visits_to_and_spending_in_uk_regions_by_purpose %>% 
  filter(region == "Scotland",
         purpose != "All visits") %>% 
  mutate(number_of_visits = number_of_visits_thousands*1000)

purposes <- c("Holiday", "Visiting friends or relatives", "Business","Miscellaneous")
purpose_visits$purpose <- factor(purpose_visits$purpose, levels = purposes, ordered = TRUE)

purpose_visits %>% 
ggplot(aes(x = year,
           y = spending_million,
           fill = purpose)) +
  geom_bar(stat = "identity", position = "dodge") +
    labs(title = "Spending in Scotland Annually by Purpose",
      x = "Year",
      y = "Spending Million (£)") +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm")) +
  scale_fill_manual(values = purpose_colours)
```

```{r}
purpose_colours <- c("#B2BABB", "#C39BD3", "#AED6F1","#533D7E")

region_visits <- visits_to_and_spending_in_uk_regions %>% 
  filter(region_vistited == "Scotland",
         region_of_residence %in% c("Europe", "North America", "Other Countries")) %>% 
  mutate(number_of_visits = number_of_visits_thousands*1000) 

region_visits%>% 
  ggplot(aes(x = year,
           y = number_of_visits,
           fill = region_of_residence)) +
   geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma)+
    labs(title = "Annual Number of Visits to Scotland by Region",
      x = "Year",
      y = "Number of Visits") +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
       legend.key.size = unit(0.5, "cm")) +
   scale_fill_manual(values = purpose_colours)
```
```{r}
purpose_colours <- c("#B2BABB", "#C39BD3", "#AED6F1","#533D7E")

region_visits <- visits_to_and_spending_in_uk_regions %>% 
  filter(region_vistited == "Scotland",
         region_of_residence %in% c("Europe", "North America", "Other Countries")) 

region_visits%>% 
  ggplot(aes(x = year,
           y = spending_million,
           fill = region_of_residence)) +
   geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::comma)+
    labs(title = "Annual Number of Visits to Scotland by Region",
      x = "Year",
      y = "Spending Million (£)") +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
       legend.key.size = unit(0.5, "cm")) +
   scale_fill_manual(values = purpose_colours)
```

```{r}

visits_to_uk_regions_by_qtr_2018_22_numeric <- visits_to_uk_regions_by_qtr_2018_22 %>%
  mutate(year_numeric = str_extract(year, "\\d+"),  # Extract the year using regular expression
         quarter = str_extract(year, "\\d+$"))  # Extract the quarter using regular expression

# Convert to numeric format by combining year and quarter
visits_to_uk_regions_by_qtr_2018_22_numeric <- visits_to_uk_regions_by_qtr_2018_22_numeric %>%
  mutate(year_quarter = as.numeric(year_numeric) + as.numeric(quarter) / 10)

```

# Line Graph of visits 
```{r}
test <- visits_to_uk_regions_by_qtr_2018_22_numeric %>%
  filter(region_visited == "Scotland",
         !(year %in% c("2020 Qtr 2", "2020 Qtr 3", "2020 Qtr 4"))) %>% 
  select(year_quarter, number_of_visits_thousands) %>% 
  group_by(year_quarter) %>% 
  summarise(number_of_visits_thousands = sum(number_of_visits_thousands)) #%>%
  # mutate(month = case_when(
  #   quarter == "1" ~ 01,
  #   quarter == "2" ~ 04,
  #   quarter == "3" ~ 07,
  #   quarter == "4" ~ 10
  # ))#,
  #date = as.Date(paste0(year_numeric, "-", month))

  
test %>% 
  ggplot(aes(x = year_quarter,
             y = number_of_visits_thousands)) +
  geom_line()
```
```{r}

test <- visits_to_uk_regions_by_qtr_2018_22_numeric %>%
  filter(region_visited == "Scotland") %>% 
  mutate(year_quarter = ifelse(is.na(year_quarter) | year_quarter == "" | year_quarter == "NA", 0, year_quarter)) %>% 
  select(year_quarter, number_of_visits_thousands, year) %>% 
  group_by(year_quarter) %>% 
  summarise(number_of_visits_thousands = sum(number_of_visits_thousands)) %>% 
  mutate(number_of_visits = number_of_visits_thousands*1000)

  
test %>% 
  ggplot(aes(x = year_quarter,
             y = number_of_visits)) +
  geom_line(color = visit_scotland_colours[1]) +
    labs(title = "Annual Number of Visits to Scotland",
      x = "Year",
      y = "Number of Visits") +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = unique(test$year_quarter),
                     labels = unique(test$year)) +
  theme_minimal() +
  theme(
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
       legend.key.size = unit(0.5, "cm")) 
    
```
```{r}
test %>% 
distinct(year_quarter)
```

```{r}
visits_by_country_2018_22 %>% 
  filter(Year == "2022") %>% 
  arrange(desc(Total))
```

Top 5 countries UK Visits 
1)USA
2)France
3) Republic of Ireland
4) Germany 
5) Spain 

```{r}
spend_by_country_2018_22 %>% 
  filter(years == "2022") %>% 
  arrange(desc(spending_million))
```

1) USA
2) France
3) Germany
4) Australia
5) Spain
